---
title: "How will new additions UCLA and USC fare in the old school Big Ten Conference?"
author: "Ryan Kenney"
date: '2022-11-06'
categories: [B1G, code, analysis]
image: ""
format: 
  html:
    code-fold: true
---

The Big Ten is an old-school football conference built on close, hard-nosed games full of tough defense and running the damn ball.

Or, at least that's what stereotypes say within college football circles. But, there does lie merit behind these stereotypes. That begs the question of how UCLA and USC, who are slated to join the Big Ten in 2024, will fare in conference play.

Entering the 2022 college football season, the Big Ten had the lowest average total points per conference game (52.5) of any FBS conference since 2014. Meanwhile, the Pac-12 averaged a full touchdown and extra point more (59.5), coming in as the second-highest scoring power five conference on average.

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggbump)
library(ggrepel)
library(cowplot)

defense14 <- read_csv("defense14.csv")
defense15 <- read_csv("defense15.csv")
defense16 <- read_csv("defense16.csv")
defense17 <- read_csv("defense17.csv")
defense18 <- read_csv("defense18.csv")
defense19 <- read_csv("defense19.csv")
defense20 <- read_csv("defense20.csv")
defense21 <- read_csv("defense21.csv")
defense22 <- read_csv("defense22.csv")

teamdefense14 <- defense14 %>% 
  mutate(
    Season = "2014"
  )

teamdefense15 <- defense15 %>% 
  mutate(
    Season = "2015"
  )

teamdefense16 <- defense16 %>% 
  mutate(
    Season = "2016"
  )

teamdefense17 <- defense17 %>% 
  mutate(
    Season = "2017"
  )

teamdefense18 <- defense18 %>% 
  mutate(
    Season = "2018"
  )

teamdefense19 <- defense19 %>% 
  mutate(
    Season = "2019"
  )

teamdefense20 <- defense20 %>% 
  mutate(
    Season = "2020"
  )

teamdefense21 <- defense21 %>% 
  mutate(
    Season = "2021"
  )

teamdefense22 <- defense22 %>% 
  mutate(
    Season = "2022"
  )

teamdefense <- bind_rows(teamdefense14, teamdefense15, teamdefense16, teamdefense17, teamdefense18, teamdefense19, teamdefense20, teamdefense21, teamdefense22)

bigten22 <- read_csv("bigten22.csv")

bigtenoffense22 <- read_csv("bigtenoffense22.csv")
bigtendefense22 <- read_csv("bigtendefense22.csv")

```

More text.

```{r}
ggplot() + 
  geom_point(
    data=bigten22, 
    aes(x=SRS, y=SOS, size=W), 
    color="grey", 
    alpha=.3) + 
  geom_point(
    data=bigten22, 
    aes(x=SRS, y=SOS, size=W), 
    color="red") + 
  geom_vline(xintercept = 7.1025) + 
  geom_hline(yintercept = 0.97125) +
  geom_text(aes(x=20.5, y=5.5, label="Good SRS, Good SOS"), color="blue") +
  geom_text(aes(x=-4.5, y=-2.8, label="Bad SRS, Bad SOS"), color="blue") +
  geom_text_repel(
    data=bigten22, 
    aes(x=SRS, y=SOS, label=Team)
  ) +
  
   labs(
    x="Simple Rating System (SRS)", 
    y="Strength of Schedule (SOS)", 
    title="USC", 
    subtitle="USC is winning games despite having a harder strength of schedule than any Big Ten team.", 
    caption="Source: Sports Reference | By Ryan Kenney"
    ) +
  
  theme_light() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 8),
    plot.subtitle = element_text(size = 10),
    panel.grid.minor = element_blank()
  )
```

```{r, fig.width=6}
bigten <- c("Illinois", "Indiana", "Iowa", "Maryland", "Michigan State", "Michigan", "Minnesota", "Nebraska", "Northwestern", "Ohio State", "Penn State", "Purdue", "Rutgers", "Wisconsin", "UCLA", "USC")

bigtendefense <- teamdefense %>% filter(School %in% bigten)

bigtenrank <- bigtendefense %>% group_by(Season) %>% 
  mutate(Rank = dense_rank(Pts))

ggplot() + 
  geom_bump(data=bigtenrank, aes(x=as.numeric(Season), y=Rank, color=School)) +
  
  geom_point(data=bigtenrank, aes(x=as.numeric(Season), y=Rank, color=School), size = 2) +
  
  geom_text(data = bigtenrank %>% filter(as.numeric(Season) == 2014), aes(x = as.numeric(Season) - .05, y=Rank, label = School), size = 2, hjust = 1) +
  geom_text(data = bigtenrank %>% filter(as.numeric(Season) == 2022), aes(x = as.numeric(Season) + .05, y=Rank, label = School), size = 2, hjust = 0) +
  
  
  scale_color_manual(values = c("#000000","#000000", "#000000", "#000000", "#000000", "#000000","#000000", "#000000", "#000000", "#000000", "#000000", "#000000", "#000000","#2D68C4","#990000","#000000")) +
  
  scale_x_continuous(breaks=c(2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022)) +
  
  scale_y_reverse(breaks=c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16)) +
  
  labs(title = "USC and UCLA on defense", subtitle="subtitle goes here", y="Rank", x="Season") +
  
theme_light()+
 theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 8), 
    plot.subtitle = element_text(size=10), 
    panel.grid.minor = element_blank()
  )
```

More text here.

```{r}
uscOff <- bigtenoffense22 %>% filter(School == "USC")
uclaOff <- bigtenoffense22 %>% filter(School == "UCLA")

bar1 <- ggplot() + 
  geom_bar(data=bigtenoffense22, aes(x=reorder(School, RushYds), weight=RushYds), fill="grey") +
  geom_bar(data=uclaOff, aes(x=reorder(School, RushYds), weight=RushYds), fill="#2D68C4") +
  geom_bar(data=uscOff, aes(x=reorder(School, RushYds), weight=RushYds), fill="#990000") +
  coord_flip() + 
  labs(title="The B1G's top passing and rushing offenses", subtitle = "Ohio State, Michigan, Michigan State and Nebraska appear on both lists.",  x="", y="Rush yards per game", caption = "") + 
  theme_minimal() + 
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 8), 
    plot.subtitle = element_text(size=10), 
    panel.grid.minor = element_blank(),
    plot.title.position = "plot"
    )

uscDef <- bigtendefense22 %>% filter(School == "USC")
uclaDef <- bigtendefense22 %>% filter(School == "UCLA")

bar2 <- ggplot() + 
  geom_bar(data=bigtendefense22, aes(x=reorder(School, OppRushYds), weight=OppRushYds), fill="grey") +
  geom_bar(data=uclaDef, aes(x=reorder(School, OppRushYds), weight=OppRushYds), fill="#2D68C4") +
  geom_bar(data=uscDef, aes(x=reorder(School, OppRushYds), weight=OppRushYds), fill="#990000") +
  coord_flip() +  
  labs(title = "", subtitle= "", x="", y="Opponent rush yards per game",  caption="Source: Sports Reference | By Ryan Kenney") + 
  theme_minimal() + 
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 8), 
    plot.subtitle = element_text(size=10), 
    panel.grid.minor = element_blank(),
    plot.title.position = "plot"
    )

plot_grid(bar1, bar2) 
```
